\documentclass[12pt]{exam}
\usepackage[utf8]{inputenc}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{multicol}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{graphicx}



\usepackage{tikz}
\usetikzlibrary{plotmarks, chains}
\usetikzlibrary{fit,positioning,arrows,automata}

\newcommand{\class}{ML \& Statistics I}
\newcommand{\term}{Winter 2017}
\newcommand{\examnum}{Exam 1}
\newcommand{\examdate}{09-Nov-2017}
\newcommand{\timelimit}{45 Minutes}

\pagestyle{head}
\firstpageheader{}{}{}
\runningheader{\class}{\examnum\ - Page \thepage\ of \numpages}{\examdate}
\runningheadrule


\begin{document}

\noindent
\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} r @{\extracolsep{6pt}} l}
\textbf{\class} & \textbf{Name:} & \makebox[2in]{\hrulefill}\\
\textbf{\term} &&\\
\textbf{\examnum} &&\\
\textbf{\examdate} &&\\
\textbf{Time Limit: \timelimit} & Teaching Assistant & \makebox[2in]{\hrulefill}
\end{tabular*}\\
\rule[2ex]{\textwidth}{2pt}

This exam contains \numpages\ pages (including this cover page) and \numquestions\ questions.\\
Total of points is \numpoints.

This test is to design the course. A similar level test will be designed after the course. Rest of introduction. Rest of introduction. Rest of introduction. Rest of introduction. Rest of introduction. All the best. 


\begin{center}
Grade Table (for coordinator's use only)\\
\addpoints
\gradetable[v][questions]
\end{center}

\noindent
\rule[2ex]{\textwidth}{2pt}
%*******************************************************
%*******************************************************
%*******************************************************
\begin{questions}

\question[3] You are working on a particular learning task and cross-validation experiments indicate that your SVM is overfitting. Name three action that can help decrease overfitting in an SVM.\\
\textbf{Answer}\\
The following are the possible ways:- \\
\begin{itemize}
    \item Decrease C
    \item Use Less expressive kernel (e.g. smaller degree polynomial)
    \item Get more training Data
\end{itemize}
\addpoints
%*******************************************************
%*******************************************************

%*******************************************************
%*******************************************************
\question[4] You have an application problem for which you need to decide whether to use a Two-Layer Neural Net or Support Vector Machine with a non-linear kernel. What would be your arguments in favor of the Net, and arguments in favor of SVM.\\
\textbf{Answer} \\
Pro SVM:
\begin{itemize}
    \item no local optima
    \item have to pick fewer parameters (e.g. layers, number of nodes) than in net
    \item easier to get estimate of generalization error, since number of support vectors in an upper bound on leave-one-out error
\end{itemize}
Pro net:
\begin{itemize}
    \item network structure allows modelling of prior knowledge
    \item can control size of the net, while size of SVM classifier grows with the number of training examples
\end{itemize}
%*******************************************************
%*******************************************************
\question[1] In the multiple linear regression model, the adjusted $R^2, \overline R^2 $ : \\
\begin{oneparchoices}
\choice cannot be negative \\
\choice will never be greater that the regression $R^2$ \\
\choice equals the square of the correlation coefficient r \\
\choice cannot decrease when an additional explanatory variable is added \\
\end{oneparchoices}
\textbf{Answer} Option B
%*******************************************************
%*******************************************************
\question[2] The following linear hypothesis can be tested using the $F$-test with the exception of \\
\begin{oneparchoices}
\choice $\beta_2$ and $\beta_3 =  \frac{\beta_4}{\beta_5}$ \\
\choice $\beta_2 = 0$ \\
\choice $\beta_1 + \beta_2 =  1$ and $\beta_3 = -2 \beta_4$ \\
\choice $\beta_0 = \beta_1$ and $\beta_1 = 0$ \\
\end{oneparchoices}
\textbf{Answer}. Option A is the exception.
%*******************************************************
%*******************************************************
\question[12] Mark box if true.
\addpoints
\begin{checkboxes}
\choice AdaBoost will eventually reach zero training error, regardless of the type of weak classifier it uses, provided enough weak classifiers have been combined. 
\textbf{False!}\\
If the data is not separable by a linear combination of the weak classifiers, AdaBoost can’t achieve zero training error. 
\choice If the outcome variable is quantitative and all explanatory variables take value 0 or 1, a logistic regression model is most appropriate.
\textbf{False!}\\
False. Logistic regression is only appropriate when the outcome variable is 0/1.
\choice Cross validation will guarantee that our model will not overfit. 
\textbf{False!}
\choice Logistic regression is a supervised machine learning algorithm.
\textbf{True!}
\choice To analyze the performance of Logistic Regression, we use AIC. We prefer a model with minimum AIC value. 
\textbf{True!}
\choice Standardization of features is required before training a Logistic Regression. 
\textbf{False!}
\choice Both LASSO and Ridge is used for variable selection \\ 
\textbf{False!}\\ Only LASSO is used
\choice probability density function's value can be greater than 1.
\textbf{True!}
\choice Exponential Distribution and Chi Square distribution are special cases of Gamma distribution.
\textbf{True!}
\choice Student's t-distribution has a thicker ($\geq$) tail than Standardized Normal distribution. 
\textbf{True!}
\choice Uniform distribution is a discrete distribution.
\textbf{False!}
\choice In PCA, the first princial component, is given by the linear combination of the original variables with the minimum possible variance.
\textbf{False!}
\choice Learning the parameters for a neural network is a convex optimization problem.
\textbf{False!}\\ It is non-convex. Easy to get, stuck in a local minima. Learning for logistic regression is a convex optimization problem. We can always find the globally optimal solution. 
\choice KNN handles non-linearly separable classes much better than logistic regression.
\textbf{True!}\\
\end{checkboxes}



{%
% changing choice items style locally
\renewcommand*\thechoice{\arabic{choice}} 
\renewcommand*\choicelabel{\thechoice)}
%
%*******************************************************
%*******************************************************
\question[1] Given the following data samples (square and triangle mean two classes), which one(s) of the following kernels can we use in SVM to separate the two classes \\
\begin{figure}[h]
\centering
\begin{tikzpicture}
\draw [very thin, lightgray] (0,0) grid (4,4);
\draw [cyan] plot [only marks, mark=square*] coordinates {(3.5,3.1) (2.2,3.33) (2.5,2.7) (0,0.5) (0.5,0.9)(0.7,0.2)(1.5,0.4)};
\draw [orange] plot [only marks, mark size=2.5, mark=*] coordinates {(0,3.5) (0.5,3.1) (0.7,2.9) (1,2.5) (1,3.0) (3,1) (3.5,0.9)(3.7,0.8)(3.5,0.5)};
\end{tikzpicture}
\caption{SVM Kernels}
\end{figure}
\begin{multicols}{2}
\begin{choices}
\choice Linear kernel. \textbf{False}
\choice Polynomial kernel. \textbf{True}
\choice Gaussian RBF (radial basis function) kernel. \textbf{True}
\choice None of the above. \textbf{False}
\end{choices}
\end{multicols}
}%

%*******************************************************
%*******************************************************
{%
% changing choice items style locally
\renewcommand*\thechoice{\arabic{choice}} 
\renewcommand*\choicelabel{\thechoice)}
%
\question[1] The parameters to be estimated in the simple linear regression model $Y = \alpha + \beta x + \epsilon \quad \epsilon \sim N(0,\sigma)$ are :\\
\begin{multicols}{2}
\begin{choices}
\choice $\alpha , \beta, \sigma$ . \textbf{True}
\choice $\alpha , \beta, \epsilon$. \textbf{False}
\choice $a , b, s$. \textbf{False}
\choice $\epsilon , 0, \sigma$. \textbf{False}
\end{choices}
\end{multicols}
}%
%*******************************************************
%*********************************************
\question[5]
In the context of classification, explain the ROC Curve \& Confusion matrix.
\makeemptybox{3in}
%*******************************************************
%*********************************************
\question[5] Suppose you want to build a decision tree for a problem. In the dataset, there are two classes with 150 examples in the + class and 50 examples in the - class.\\
Recall the definitions of information gain and entropy:
\begin{align*}
Entropy(C) &= H(C) = \sum_c - P(C=c)log_2P(C=c) \\
Gain(C,A) &= H(C) - \sum_{v \in Values(A)} P(A=v) H(C|A=v)
\end{align*}
What is the entropy of the class variable (you can leave this in terms of logs).\\
\textbf{Answer}. 
\begin{equation}
    H(C) = -\frac{150}{150+50} log \frac{150}{150+50} -\frac{50}{150+50} log \frac{50}{150+50}
\end{equation}

\addpoints
%*******************************************************
%*******************************************************
\question[5]
How does classification differ from clustering? \\
Explain the difference between K-nearest neighbor classification \& k-means clustering in terms of their algorithm steps.
\makeemptybox{3in}
%*******************************************************
%*******************************************************
\question[3]Consider a Markov chain whose transition diagram is as given. 
\noaddpoints % to omit double points count
\begin{parts}
\part[1] Which (if any) states are absorbing?
\part[1] Find the communicating classes.
\part[1] Is the chain irreducible?
\end{parts}
\addpoints
\begin{figure}[!h]
\centering
\begin{tikzpicture}[
mynode/.style={
  draw,
  minimum size=1em
  },
every loop/.append style={-latex},  
start chain=going right  
]
\foreach \Value in {1,...,5}
  \node[mynode,on chain] (s\Value) {$S_{\Value}$};
\path[-latex]
  (s2) edge[bend right] node[auto,swap,font=\small] {$0.7$} (s1)
  (s2) edge[bend right] node[auto,swap,font=\small] {$0.3$} (s3)
  (s3) edge[bend right] node[auto,swap,font=\small] {$0.5$} (s2)
  (s3) edge[bend right] node[auto,swap,font=\small] {$0.5$} (s4)
  (s4) edge[bend right] node[auto,swap,font=\small] {$0.65$} (s3)
  (s4) edge[bend right] node[auto,swap,font=\small] {$0.35$} (s5)
  (s1) edge[loop left] node[left,font=\small] {$1$} (s1)
  (s5) edge[loop right] node[right,font=\small] {$1$} (s5);
\end{tikzpicture}
\caption{Markov Chain}
\end{figure}
%*******************************************************
%*********************************************
\question[4]
Explain strong stationarity \& weak stationarity
\makeemptybox{\fill}
%*******************************************************
%*********************************************
\question[4]
How is $p, q$ decided in $ARMA(p,q)$ model? 
\makeemptybox{\fill}
%\newpage
%*******************************************************
%*******************************************************
\question[5]
Ridge Regression, Lasso Regression, and Elastic Net are used to implement three different ways to constrain the weights to achieve regularization for a linear model. What are those ways?
%\fillwithlines{\fill}

%\newpage
%*******************************************************
%*******************************************************
\question[5]
The \textit{perceptron} is one of the simplest Artificial Neural Network architecture. Explain the algorithm and how it works.

\fillwithdottedlines{8em}
%*******************************************************
%*******************************************************
\question[6] Consider the following HMM: 
\begin{figure}[!htb]
\begin{tikzpicture}
\clip (-2.5,-2.5) rectangle (8.5,2.5);
\tikzstyle{main}=[circle, minimum size = 5mm, thick, draw =black!80, node distance = 10mm]
\tikzstyle{connect}=[-latex, thick]
\tikzstyle{box}=[rectangle, draw=black!100]
  \node[box,draw=white!100] (Latent) {\textbf{Latent}};
  \node[main] (L1) [right=of Latent] {$X_1$};
  \node[main] (L2) [right=of L1] {$X_2$};
  %\node[main] (L3) [right=of L2] ;
  %\node[main] (Lt) [right=of L3] ;
  \node[box,draw=white!100] (Observed) [below=of Latent] {\textbf{Observed}};
    \node[main,fill=black!10] (O1) [right=of Observed,below=of L1] {$O_1$};
  \node[main,fill=black!10] (O2) [right=of O1,below=of L2] {$O_2$};
 % \node[main,fill=black!10] (O3) [right=of O2,below=of L3] {$O_3$};
  %\node[main,fill=black!10] (Ot) [right=of O3,below=of Lt] {$O_t$};
  \path (L3) -- node[auto=false]{\ldots} (Lt);
  \path (L1) edge [connect] (L2) ;
        %(L2) edge [connect] (L3)
        %(L3) -- node[auto=false]{\ldots} (Lt);
  %\path (O1) edge [connect] (O2) ;
        %(O2) edge [connect] (O3)
        %(O3) -- node[auto=false]{\ldots} (Ot);
  \path (L1) edge [connect] (O1);
  \path (L2) edge [connect] (O2);
  %\path (L3) edge [connect] (O3);
  %\path (Lt) edge [connect] (Ot);
  \draw[dashed]  [below=of L1,above=of O1];

\path (Latent) -- (Observed) coordinate[midway](l44);
\node (l43) [left=of l44] {};
%\path (Lt) -- (Ot) coordinate[midway](l444);
\node (l433) [right=of l444] {};
\draw[dashed,thick] (l43) to ++(0:150);

\end{tikzpicture}
\end{figure}





\begin{table}[!htb]
    \caption{HMM}
    \begin{minipage}{.2\linewidth}
      \caption{}
      \centering
        \begin{tabular}{ |c|c| } 
 \hline
 $X_1$ & $P(X_1)$  \\ 
 \hline
 0 & 0.3  \\ 
 1 & 0.7  \\ 
 \hline
\end{tabular}
    \end{minipage}%
     \begin{minipage}{.4\linewidth}
      \centering
        \caption{}
        \begin{tabular}{ |c|c|c| } 
 \hline
 $X_t$ & $P(X_{t+1})$ &  $P(X_{t+1}|X_t)$\\ 
 \hline
 0 & 0 & 0.4  \\ 
 0 & 1 & 0.6  \\ 
 \hline
 1 & 0 & 0.8  \\ 
 1 & 1 & 0.2  \\
 \hline
\end{tabular}
    \end{minipage}
    \begin{minipage}{.4\linewidth}
      \centering
        \caption{}
        \begin{tabular}{ |c|c|c| } 
 \hline
 $X_t$ & $P(X_{t+1})$ &  $P(X_{t+1}|X_t)$\\ 
 \hline
 0 & A & 0.9  \\ 
 0 & B & 0.1  \\ 
 \hline
 1 & A & 0.5  \\ 
 1 & B & 0.5  \\
 \hline
\end{tabular}
    \end{minipage} 
\end{table}
\noaddpoints % to omit double points count
\begin{parts}
\part[3] Use the forward algorithm to compute the probability distribution $P(X_2, O_1 = A, O_2 = B)$. Show your work. You do not need to evaluate arithmetic expressions.
\makeemptybox{1.5in}
\part[3] Compute the probability distribution $P(X_1 = 1| O_1 = A, O_2 = B)$. Show your work.
\makeemptybox{1.5in}
\end{parts} 

\addpoints
%*******************************************************
%*******************************************************
\question[4]
Explain the following equation (in terms of multivariate linear regression):\\
\begin{equation}
\hat{\beta} = (X^TX)^{-1}X^TY
\end{equation}
Give numbers for the dimensions of the matrix as well
\makeemptybox{3in}
%*******************************************************
%*******************************************************
\end{questions}

\end{document}
